Chaotic mixing of tidal debris

### 2015-06-22 ms.tex
so many words

% Context
The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape.
Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable. 
An appreciable number of chaotic orbits are therefore expected in galactic halos.
It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
% Aims
Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
% Methods: 
We derp derp derp.
% Results:
We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
% Conclusions:

@@

When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}. We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion. Any transformations to angle-action coordinates must be defined locally due to the destruction of tori due to chaotic motion.

@@

The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict, however, this has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}. For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods. For galactic timescales, Arnold diffusion is not expected to be relevant, however chaotic motion across resonances can occur over short times. [the volume allowed to explore over short times is small, but could be relevant for tidal debris...]

@@

Small ensembles of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}. Ensembles of chaotic orbits behave quite differently. Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori). Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}. Certain ensembles may be trapped between resonances for long times and thus remain in a state of pseudo-coherence for long times, whereas others may rapidly diffuse \citep[see, e.g.,][]{merritt96}. This phenomenon is generic to potentials that are not globally stochastic in which regions of stability and chaos are intermingled throughout phase-space.

\todo{apw}{concluding paragraph...}

@@

This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies. Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}. We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations. \todo{apw}{Cite this paper: hunter05 for disk crossings and chaos}

@@

The MLE, $\lambda_{\rm max}$, is interpreted as a rate that sets the exponential divergence of infinitesimally close chaotic orbits. It is therefore useful to consider the corresponding $e$-folding time by inverting the rate. We normalize the Lyapunov time by the maximum orbital period of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
\begin{equation}
	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max} \, T_{\rm max}}
\end{equation}
We will use this as the prediction from the Lyapunov exponent for the timescale over which chaos should be dynamically important for a given orbit.

@@

Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$). It has been shown that using a filtered FFT makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time. The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration. This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.

This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested a version of this procedure in the \project{Python} programming language. Our implementation differs slightly from the original definition and from that used in \cite{valluri98}: we have found that using a higher order window function \citep[e.g.,][]{hunter02} of the form
\begin{equation}
	W(t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \frac{\pi t}{\inttime}\right)^p
\end{equation}
with $p=2$ allows for more reliable determination of frequencies from strongly chaotic orbits. We refer to this slight modification of the algorithm as \superfreq\ and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}. For more details about the algorithm, see Appendix~\ref{sec:naffapdx} or \cite{laskar88, laskar93, papaphilippou96}.

If an orbit is chaotic, the motion can no longer be expanded in terms of a single set of fundamental frequencies. For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time. \superfreq\ will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time. For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two separate sections of orbit. The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window. We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $k$th fundamental frequency as
\begin{equation}
	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, N} \label{eq:fdrate}
\end{equation}
where the upper index refers to which of the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit. By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
\begin{equation}
	t_\Omega = \, (\max_{a_k} \, \mathcal{R}_k)^{-1} \label{eq:fdtime}
\end{equation}
where the maximum is taken with respect to the corresponding amplitudes, $a_i$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.

For a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble. We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris. For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (number of orbital periods) over which we expect the frequencies to evolve by this amount.

@@

We generate isoenergy grids of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}). We fix $v_x = v_z = 0$, and compute $v_y$ from the energy. The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic. Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable regions of stochastic tube orbits. Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits. The major classes of stable tube orbits circulate about either the major or minor axis. This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes.

@@

We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S). The orbits were chosen at random from a single column of the orbit grid based on their frequency diffusion time, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}. We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits. The initial mass of the clusters is \todo{kvj}{2.5E4 \msun contained within what again? 35 scale radii? but this is a plummer sphere?}. For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking. We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter. Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis. A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a ``fanned'' stream on the strongly chaotic orbit (right column). Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.

\begin{table*}[ht]
\begin{center}
	\begin{tabular}{c | c c c c c c | c c }
		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
		R & 24.90 & 0 & 6.90 & 0 & 190.76 & 0 & XX & YY\\
		\midrule
		W & 24.90 & 0 & 19.70 & 0 & 137.19 & 0 & XX & YY\\
		\midrule
		S & 24.90 & 0 & 26.10 & 0 & 93.11 & 0 & XX & YY\\
		\bottomrule
		\end{tabular}
	\caption{Orbits from the $xz$ grid with a range of chaotic timescales --- R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit. Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods. \label{tbl:three-orbits}}
\end{center}
\end{table*}

It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits. We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event. For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time. The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}. If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
\begin{align}
	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
\end{align}
where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the debris, and $\|\cdot \|$ is the Euclidean norm. We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales. Figure~\ref{fig:ensembles} shows final positions of test-particle0 ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.

@@

The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles. Figure~\ref{fig:densities} shows the mean density of the ensemble particles computed at 128 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods. It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse.

Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods ---we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}. For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density at the end of integration for each. Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density. \todo{apw}{talk about the structure here, compare to freq. diff time map -- even though freq. diff time are long, still resolve structure in these maps.}

@@

As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}). Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits. The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods. The resonant structure around the mildly chaotic orbit (white regions in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic. To resolve the short-time behavior of the frequency diffusion (corresponding to motion across resonance layers) we compute the frequencies in a series of rolling windows along each orbit. We use a window with a width equal to 40 orbital periods and shift the window by half an orbital period between each calculation of the most significant frequencies --- this is the smallest window size we could choose before \superfreq\ begins to consistently fail to recover the frequencies for the strongly chaotic orbit.  Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the median value over all windows. The difference between the first window (green +) and the last window (red x) is 50 orbital periods. The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces). This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%).

@@

>>>>

% Context
%The large-scale, smooth components of galactic gravitational potentials are thought to be triaxial in shape.
%Triaxial potentials that match observational constraints and measurements of the radial profile and axis ratios of dark matter halos in cosmological simulations are almost certainly not globally integrable. 
%An appreciable number of chaotic orbits are therefore expected in galactic halos.
%It is yet unclear whether chaotic regions of phase-space will affect the density evolution of stellar tidal streams, which form in the halos of galaxies from the disruption of stellar systems.
%% Aims
%Many numerical methods exist for assessing the degree of chaos for a given orbit (e.g., the Lyapunov exponent, the frequency diffusion rate). 
%These indicators, which are local to a single orbit, often predict a timescale over which chaos will be dynamically important.
%In this paper, we aim to compare these timescales to the density evolution of the finite-volume ensembles of orbits.
%% Methods: 
%We derp derp derp.
%% Results:
%We find that mean indicators such as the Lyapunov time or the (long term) frequency diffusion rate of a given ``parent'' orbit are not good predictors for the short-term density evolution of ensembles of orbits around this parent orbit.
%Instead, we find that the variance of the fundamental frequencies of the ensemble orbits over short timescales is a stronger indicator of the density evolution.
%% Conclusions:

@@

When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep[e.g.,][]{valluri98, merritt99}. We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion. Any transformations to angle-action coordinates must be defined locally due to the destruction of tori due to chaotic motion.

@@

The rate of diffusion along stochastic layers via Arnold diffusion depends on the local resonant structure and is hard to predict. This has been done analytically for simple potentials \citep[e.g.,][]{chirikov??}. For systems with $N>2$ dof, the stochastic layers connect and form an intricate network of stochasticity known as the Arnold web; an orbit that ergodically mixes over its energy hypersurface must traverse this web, though the timescales typical for this phenomenon are many thousands of orbital periods.

For galactic timescales, Arnold diffusion is not expected to be relevant, however chaotic motion across resonances can occur over short times. If a stochastic orbit is semi-bounded by surrounding resonances, the allowed volume in which it can explore (in frequency space) is often quite small so that the orbit may appear to be approximately regular. However, this small-scale evolution may be important for tidal debris which has small intrinsic spreads in frequencies.

@@

Ensembles of orbits stuck between resonances may quickly spread to fill the allowed volume via the small-scale but short-time chaotic diffusion processes discussed above. This occurs due to phase-mixing of the fundamental frequencies of the ensemble orbits themselves --- the frequencies are evolving and thus trace out an orbit through frequency space. This is in contrast to phase-mixing of regular orbits and long-term chaotic mixing of strongly chaotic orbits.

A small ensemble of (conditionally-periodic) regular orbits will reach a fully phase-mixed state in a timescale  $\approx\sigma_\Omega^{-1}$, where $\sigma_\Omega$ is the dispersion in frequency-space of the ensemble --- each individual orbit will fill the surface of its torus ergodically, but the ensemble will only mix due to sheering from slight variations in their fundamental frequencies \citep{merritt96, helmi99}. Ensembles of chaotic orbits behave quite differently. Over long times, a chaotic orbit in an autonomous Hamiltonian will ergodically mix over the energy hypersurface (a 5D surface, compared to the 3D stable tori). Generically, a small ensemble of chaotic orbits will lose coherence much faster than for regular orbits, however analytically computing the rate and amount of spreading is very difficult even for simple potentials \cite[see, e.g.,][]{chirikov??}. Typically these rates are comparable or larger than the Arnold diffusion timescale and are thus very long.

In this work, we investigate the effect of the short-time but small-scale frequency evolution along weakly chaotic orbits and discuss how this effects our understanding of the coherence and density evolution of tidal streams.

@@

This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies. Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}. We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.

@@

The MLE, $\lambda_{\rm max}$, is interpreted as a rate that quantifies the exponential divergence of infinitesimally close chaotic orbits. It is therefore useful to consider the corresponding $e$-folding time by inverting the rate. We normalize the Lyapunov time by the maximum orbital period\footnote{The maximum of the three orbital periods corresponding to the three fundamental frequencies (see Section~\ref{sec:naff}).} of a given orbit, $T_{\rm max}$, so that the time is in units of number of orbital periods
\begin{equation}
	t_{\rm \lambda} = \frac{1}{\lambda_{\rm max} \, T_{\rm max}}
\end{equation}
We will use this as the prediction from the Lyapunov exponent for the timescale over which chaos should be dynamically important for a given orbit.

@@

Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$). It has been shown that the accuracy in determining the frequencies converges much faster using a filtered FFT compared to the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time. The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration. This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.

This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested a version of this procedure in the \project{Python} programming language. Our implementation differs slightly from the original definition and from that used in \cite{valluri98}: we have found that using a higher order window function \citep[e.g.,][]{hunter02} of the form
\begin{equation}
	W(\tau=t/\inttime) = \frac{2^p \, (p!)^2}{(2p)!} \left( 1 + \cos \pi \tau\right)^p
\end{equation}
with $p=2$ allows for more reliable determination of frequencies from strongly chaotic orbits. We refer to this slight modification of the algorithm as \superfreq\ \citep{superfreq} and the code is open-source and publicly available on \project{GitHub} at \url{https://github.com/adrn/SuperFreq}. For more details about the algorithm, see Appendix~\ref{sec:naffapdx} or \cite{laskar88, laskar93, papaphilippou96}.

If an orbit is chaotic, the motion can no longer be expanded in terms of a single set of fundamental frequencies. For a weakly chaotic orbit, the orbit may appear consistently periodic over long windows of time. \superfreq\ will pick out a set of frequencies for such orbits that correspond to the largest peaks in the power spectrum of the stochastic orbits, however these peaks will change location and amplitude with time. For more strongly chaotic orbits, the power spectrum will be quite noisy and the peak frequencies may change erratically when comparing two consecutive sections of orbit. The frequencies picked out by \superfreq\ for such orbits will therefore represent the average periodic nature of the orbit over a given integration window. We define the fractional frequency diffusion rate per orbital period, $\mathcal{R}$, in the $k$th fundamental frequency as
\begin{equation}
	\mathcal{R}_k = \frac{\Omega_{k}^{(2)} - \Omega_{k}^{(1)}}{\Omega_{k}^{(1)} \, N} \label{eq:fdrate}
\end{equation}
where the upper index refers to the two consecutive sections of orbit and $N$ is the number of orbital periods in a single section of orbit. By inverting this rate, we can compute the timescale over which we expect order-unity changes to the fundamental frequencies: the \emph{frequency diffusion time} is defined as
\begin{equation}
	t_\Omega = \, (\max_{a_k} \, \mathcal{R}_k)^{-1} \label{eq:fdtime}
\end{equation}
where the maximum is taken with respect to the corresponding amplitudes, $a_k$, of the fundamental frequency components \citep[see][]{valluri12}, and the time is in units of number of orbital periods, ${\rm T}$.

For a small ensemble of orbits (e.g., tidal debris), the relevant timescale is the time over which the change in frequencies for a single orbit is comparable to the spread of frequencies in the ensemble. We can estimate this timescale by multiplying the frequency diffusion time by a factor equal to the fractional spread in frequencies of the debris. For example, a globular cluster typically has $\approx$1\% spreads in fundamental frequencies, so by multiplying the frequency diffusion time by $f = 0.01$, we can estimate the time (in number of orbital periods) over which we expect the frequencies to evolve by this amount.

@@

We generate an isoenergy grid of initial conditions along the $xz$ ($y=0$) plane\footnote{$x$ is the major and $z$ the minor axis.} with energy (per unit mass), $E$, chosen to span a range of distances comparable to the scale radius of the potential ($E=-(397.2~{\rm km~s}^{-1})^2$ in physical units; see Table~\ref{tbl:potential}). We fix $v_x = v_z = 0$, and compute $v_y$ from the energy. The majority of the orbits generated on this grid are tube orbits, which preserve a sense of rotation about one of the principal axes of the potential and are generally centrophobic. Tube orbits are generally less stochastic than box orbits, which tend to plunge deep into the inner regions of the potential where the density profile is steep, however there are appreciable numbers of stochastic tube orbits. The major classes of stable tube orbits circulate about either the major or minor axis. This grid generates all of the major orbit classes (short-axis tubes, inner long-axis tubes, outer long-axis tubes, stochastic intermediate-axis, and box orbits), but the most numerous orbits in this grid are the short-axis and outer long-axis tubes. Thin tidal streams are thought to form along tube orbits rather than box orbits because of the faster disruption expected for stellar systems on radially plunging orbits.

@@

We start by considering three particular orbits chosen from the orbit grid of Section~\ref{sec:results1}: a regular orbit (R), a weakly chaotic orbit (W), and a strongly chaotic orbit (S). The orbits were chosen from a single column of the orbit grid based on their frequency diffusion time to sample a range of possible values, and the initial conditions and chaos diagnostics are listed in Table~\ref{tbl:three-orbits}. We use the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to run N-body simulations of globular-cluster-mass progenitor systems on these three orbits. The initial mass of the clusters is \todo{kvj}{2.5E4 \msun contained within what again? 35 scale radii? but this is a plummer sphere?}. For each simulation, the progenitor is initialized as a $10^4$ particle Plummer sphere and the background triaxial NFW potential is turned on slowly over 250 Myr to reduce artificial gravitational shocking. We start the progenitor systems at apocenter and evolve for $\approx$16 orbital periods so that each simulation finishes again at subsequent apocenter. Figure~\ref{fig:nbodysims} shows the final particle distributions rotated so that the angular momentum of the progenitor orbit is aligned with the $z$-axis. A thin stream forms on the regular orbit (left column), a more diffuse stream on the mildly chaotic orbit (middle column), and a ``fanned'' stream on the strongly chaotic orbit (right column). Given the long Lyapunov and frequency diffusion times of the mildly chaotic orbit, it is surprising that stream formed on this orbit appears to be more diffuse than the regular orbit stream.

\begin{table*}[ht]
\begin{center}
	\begin{tabular}{c | c c c c c c | c c }
		{\bf ID} & $\bs{x}$ & $\bs{y}$ & $\bs{z}$ & $\bs{v_x}$ & $\bs{v_y}$ & $\bs{v_z}$ & $\bs{t_\lambda}$ & $\bs{t_\Omega}$ \\\toprule
		R & 29.10 & 0 & 5.10 & 0 & 168.07 & 0 & $>10^3$ & $>10^8$\\
		\midrule
		W & 29.10 & 0 & 17.70 & 0 & 125.49 & 0 & $\approx55$ & $\approx8 \times 10^4$\\
		\midrule
		S & 29.10 & 0 & 22.10 & 0 & 99.45 & 0 & $\approx3.5$ & $\approx6 \times 10^3$\\
		\bottomrule
		\end{tabular}
	\caption{Orbits from the $xz$ grid with a range of chaotic timescales --- R is a regular orbit, W a weakly chaotic orbit, and S a strongly chaotic orbit. Positions ($x$, $y$, $z$) are given in kpc, velocities ($v_x$, $v_y$, $v_z$) in km~s$^{-1}$, and times ($t_\lambda$, $t_\Omega$) in number of orbital periods. Recall that the frequency diffusion time, $t_\Omega$, is the time over which we expect order-unity changes in the fundamental frequencies, hence why the timescales appear quite long. \label{tbl:three-orbits}}
\end{center}
\end{table*}

It is computationally intractable to run full N-body simulations for the large grid of orbital initial conditions of the previous section and we therefore take a simplified approach for studying how finite-volume debris spreads along each of these orbits. We instead consider small ensembles of particles meant to represent debris disrupted from a single tidal disruption event. For a given set of orbital initial conditions --- the ``parent'' orbit --- we find the phase-space position of the nearest pericenter by integrating forward for a short time until a minimum is found, initialize a small ensemble of test particle orbits around this position, then follow the orbits of all test particles for a specified integration time. The physical scale of the ensemble is set by the tidal radius in position and the velocity scale in velocity and are therefore set by the mass of the progenitor \citep[e.g.,][]{johnston98, apw14}. If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and the magnitude of the offsets are assumed to be Normally distributed away from the parent orbit:
\begin{align}
	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
	r_{\rm tide} &= \|\bs{x}_0\| \left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} \\
	\sigma_v &= \|\bs{v}_0\|\left(\frac{m}{3M(<\|\bs{x}_0\|)}\right)^{1/3} 
\end{align}
where $M(<r)$ is the mass enclosed of the host potential within radius $r$, $m$ is the mass scale of the ``progenitor,'' and $\|\cdot \|$ is the Euclidean norm. We take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass of the host potential to estimate the above debris scales. Figure~\ref{fig:ensembles} shows final positions of test-particle ensembles initialized around the three orbits described above and evolved for 16 orbital periods (same as the N-body simulations); the morphology of the ensembles is visually similar to the ``oldest'' (first stripped) debris in the N-body simulations.

@@

\todo{apw}{These paragraphs need work, but waiting for results from the cluster...}
The density distributions do not change shape dramatically, even for the strongly chaotic ensemble; we therefore use the mean density as a metric for studying the density evolution of orbit ensembles. Figure~\ref{fig:densities} shows the mean density of the ensemble particles computed at 128 evenly-spaced intervals from the initial ensemble distribution to the distribution after 16 orbital periods. It is clear that the density of the regular orbit ensemble evolves slower than the weakly chaotic ensemble, and the strongly chaotic ensemble quickly becomes diffuse.

Motivated by the noticeable discrepancy between the regular and weakly chaotic orbit ensembles --- even though the chaotic timescale of the weakly chaotic ensemble parent orbit is thousands of orbital periods --- we compute the final mean density for orbit ensembles generated around each orbit in the grid described in Section~\ref{sec:results1}. For all ensembles, we integrate the orbits for 16 (parent) orbital periods and store the mean density at the end of integration for each. Figure~\ref{fig:ensemblemap-meandensity} again shows the grid of initial conditions (Section~\ref{sec:results1}), but now the greyscale indicates the final mean density. \todo{apw}{talk about the structure here, compare to freq. diff time map -- even though freq. diff time are long, can still resolve structure in these maps.}

@@

As a demonstration of the small-scale frequency modulation, we consider again the three orbits of Section~\ref{sec:results2} (initial conditions are listed in Table~\ref{tbl:three-orbits}). Figure~\ref{fig:fdiff-zooms} shows zoomed regions of the frequency diffusion time map of Figure~\ref{fig:freqdiff} centered on the three orbits --- the pixel positions are now shown relative to the three parent orbits. The frequency diffusion times shown in these grids were computed with two consecutive windows of length 40 orbital periods. The resonant structure around the mildly chaotic orbit (white lines in panel W) bound the orbit to a small region, whereas for the strongly chaotic orbit (S), the entire region surrounding the orbit is strongly chaotic. To resolve the short-time behavior of the frequency diffusion (corresponding to motion across or around resonance layers) we compute the frequencies in a series of rolling windows along each orbit. We use a window with a width equal to 40 orbital periods and shift the window by an orbital period between each calculation of the most significant frequencies --- again, this is the smallest window size we could choose before \superfreq\ begins to occasionally fail to recover the frequencies for the strongly chaotic orbit.  Figure~\ref{fig:three-orbits-freqs} shows plots of the frequency ratios as computed in each window of time --- plotted are the percent deviations of the frequency ratios from the initial value. The difference between the first window (green +) and the last window (red x) is 50 orbital periods. The frequencies of the weakly chaotic orbit (panel labeled $W$) evolve quickly but are bounded to a small volume with fractional size $\approx$0.3\% (presumably by nearby resonant surfaces). This is comparable to the frequency spread in globular cluster tidal debris (0.1--0.5\%). %\todo{apw}{check numbers in previous 2 sentences}

@@

\subsection{The progenitor mass scale}

In this work, we have only considered low-mass progenitor systems such as globular clusters because the intrinsic spreads in fundamental frequencies are small (0.1--0.5\%). Small changes to the frequencies of the orbits of tidal debris disrupted from these progenitors due to chaotic processes will therefore cause observable changes to the real-space morphology of the debris. For more massive progenitor systems, the typical size and velocity dispersion of the debris will be larger and thus the debris morphology will be less sensitive to small changes in orbital properties. The mass-scale of the debris that will display enhanced density evolution depends on the magnitude of weak chaos, which depends on the orbit structure of a given potential. In potentials with more significant chaos, debris disrupted from more massive progenitor systems may also display stream-fanning.
